Bow shocks, bow waves, and dust waves. III. Diagnostics

### 2019-03-31 papers/Papers1234/sec-observational-diagnostics.tex
Finished second draft of Paper III

A fundamental parameter is the optical depth, \(\tau\), of the bow
shell to UV radiation, which determines what fraction of the stellar
photon momentum is available to support the shell (see
\S~2.1 of Paper~I).  But the same photons also heat the
dust grains in the bow, which re-radiate that energy predominantly at
mid-infrared wavelengths (roughly \SIrange{10}{100}{\um}) with
luminosity \(L\IR\).  Assuming that Ly\(\alpha\) and mechanical
heating of the dust shell is negligible (see
\S~\ref{sec:unimp-other-heat}) and that the emitting shell subtends a
solid angle \(\Omega\), as seen from the star, then the optical depth
can be estimated as
\begin{equation}
  \label{eq:tau-empirical}
  \tau = -\ln \left( 1 - \frac{4\pi}{\Omega} \frac{L\IR}{L_*} \right)
  \approx \frac{2 L\IR}{L_*} \ ,
\end{equation}
where the last approximate equality holds if \(\tau \ll 1\) and the shell
emission covers one hemisphere.  Note that the \(\tau\) of Paper~I is not
exactly the same as the \(\tau\) of equation~\eqref{eq:tau-empirical},
but is larger by a factor of
\(Q_P / Q_{\text{abs}} = 1 + \varpi (1 - g)/(1 - \varpi)\), where
\(\varpi\) is the grain albedo and \(g\) the scattering asymmetry.
This is plotted in Figure~6a of Paper~II for Cloudy's default ISM
grain mixture (see \S~\ref{sec:grain-temp-emiss} below), where it can
be seen that \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at
EUV/FUV wavelengths (\(\lambda = \text{\SIrange{0.05}{0.2}{\um}}\)), so the
correction is small.

A second important parameter is the thermal plus magnetic pressure in
the shocked shell, which is doubly useful since in a steady state it
is equal to \emph{both} the internal supporting pressure (wind ram
pressure plus absorbed stellar radiation) \emph{and} the external
confining pressure (ram pressure of ambient stream).  The shell
pressure is not given directly by the observations, but can be
determined by the following three steps:
\begin{enumerate}[P1.]
\item \label{P1} The shell mass column (\si{g.cm^{-2}}) can be
  estimated from the optical depth by assuming an effective UV
  opacity: \(\Sigma\shell = \tau / \kappa\)
\item \label{P2} The shell density (\si{g.cm^{-3}}) can be found from
  the mass column if the shell thickness is known:
  \(\rho\shell = \Sigma / h\shell\).  In the absence of other information, a
  fixed fraction of the shell radius can be used.  In particular, we
  normalize by a typical value of one~quarter the star--apex distance:
  \(h_{1/4} = h\shell / (0.25 R_0)\).  This corresponds to a Mach
  number \(\M_0 = \surd 3\) if the stream shock is radiative, or
  \(\M_0 \gg 1\) if non-radiative (see \S~3.2 of Paper~I). Further
  discussion is given in \S~\ref{app:bow-shock-data} below. 
\item \label{P3} Finally, the pressure (\si{dyne.cm^{-2}}) follows by
  assuming values for the sound speed and Alfvén speed:
  \(P\shell = \rho\shell (\sound^2 + \frac12 v\alfven^2) \).
\end{enumerate}
It is natural to normalize this pressure to the stellar radiation
pressure at the shell, so we define a shell momentum efficiency
\newcommand\pc{\ensuremath{_{\text{pc}}}}
\begin{equation}
  \label{eq:eta-shell}
  \eta\shell \equiv \frac{P\shell}{P\rad}
  = \frac{4\pi R_0^2\, (\sound^2 + \frac12 v\alfven^2)\, \tau}{L_*\, \kappa\, h\shell}
  \approx 245 \frac{R\pc \, T_4 \, \tau}{L_4 \, \kappa_{600} \, h_{1/4}} \ , 
\end{equation}
where in the last step we have assumed ionized gas with negligible
magnetic support (\(v\alfven \ll \sound\)) and written the stellar
luminosity and shell parameters in terms of typical values, which we
summarize below.  Note that the shell momentum efficiency is simply
the reciprocal of the radiation parameter from Paper~II's equation~(23):
\(\eta\shell = \Xi\shell^{-1}\), which provides yet a third use for
\(\eta\shell\), since \(\Xi\) is paramount in determining whether the
grains and gas remain well-coupled (see \S~4.4 of Paper~II).

@@

\begin{table}
  \centering
  \caption[Observational]{Key observational parameters for star/bow systems}
  \label{tab:observations}
  \begin{tabular}{l S S S}
    \toprule
    Star & {\(L_* / \si{L_\odot}\)} & {\(L_{\text{IR}} / \si{L_\odot}\)} & {\(R_0 / \si{pc}\)} \\
    \midrule
    \thD & 2.95e4 & 620 & 0.003 \\
    LP~Ori & 1600 & 240 & 0.01 \\
    \(\sigma\)~Ori & 6e4 & 15 & 0.12 \\[\smallskipamount]
    K18 Sources & \numrange{1.4e4}{8.7e5} & \numrange{8}{2800} & \numrange{0.02}{1.35} \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The UV optical depth of the bow shell is obtained
(eq.~[\ref{eq:tau-empirical}]) from the ratio of infrared shell
luminosity to stellar luminosity.  The inverse of this ratio is given
in Table~5 of K17, but we choose to re-derive the values since the
spectral classification of some of the sources was revised between K17
and K18.  Although K17 found the total shell fluxes from fitting dust
emission models to the observed SEDs, we adopt the simpler approach of
taking a weighted sum of the flux densities \(F_\nu\) (in \si{Jy}) in
three mid-infrared bands:
\begin{multline}
  \label{eq:total-ir-flux}
  F_{\text{IR}}  \approx \bigl[  2.4\,(F_8 \text{ or } F_{12})
    + 1.6\,(F_{22} \text{ or } F_{24})  \\
  + 0.51\,F_{70}\bigr]
  \,\times \SI{e-10}{erg.s^{-1}.cm^{-2}} \ ,
\end{multline}
where \(F_8\) is Spitzer IRAC \SI{8.0}{\um}, \(F_{24}\) is Spitzer
MIPS \SI{23.7}{\um}, \(F_{12}\) and \(F_{22}\) are WISE bands 3 and 4,
and \(F_{70}\) is Herschel PACS \SI{70}{\um}.  The weights are chosen
so that the integral \(\int_0^\infty F_\nu \,d\nu\) is approximated by the sum
\(\Sigma_k F_k\, \Delta\nu_k\), under the assumption that fluxes in shorter (e.g.,
IRAC \SI{5.8}{\um}) and longer (e.g., PACS \SI{150}{\um}) wavebands
are negligible.  Shell fluxes are converted to luminosities using the
assumed distance to each source, and stellar luminosities are taken
directly from K18 Table~2, based on spectroscopic classification and
the calibrations of gravity and effective temperature from
\citet{Martins:2005a}.

@@

The fundamental observational quantities that go into determining
\(\tau\) and \(\eta\shell\) for each source are distance, \(D\); stellar
luminosity, \(L_*\); total infrared flux, \(F_{\text{IR}}\); and bow
angular apex distance, \(\theta\).  From these, the shell radius and
infrared luminosity are found as \(R_0 = \theta D\) and
\(L_{\text{IR}} = 4\pi D^2 F_{\text{IR}}\).  Rather than clutter the
diagram with error bars, we instead show the sensitivity to
observational errors in the lower-right box, where each arrow
corresponds to a factor of two increase (0.3~dex) in each quantity:
\(D\), \(L_*\), \(F_{\text{IR}}\), and \(\theta\).  We now calculate
uncertainty estimates for individual observational quantities that are
used in deriving not only \(\tau\) and \(\eta\shell\), but also mass-loss
rates, as discussed later in \S~\ref{sec:stellar-wind-mass}.

@@

Assuming that the uncertainty in each observational quantity is
independent, we can now combine them using the techniques described in
Appendix~\ref{sec:comb-uncert-covar} to find the \(\pm 1~\sigma\) error
ellipse, shown in blue in the figure.  It can be seen that
observational uncertainties in \(\tau\) and \(\eta\shell\) are highly
correlated: the dispersion is \SI{0.7}{dex} in the product
\(\eta\shell \tau\) but only \SI{0.16}{dex} in the ratio
\(\eta\shell/\tau\), with stellar luminosity errors dominating in both
cases.  Observational uncertainties are therefore relatively
unimportant in determining whether a given source is wind-driven or
radiation-driven, which depends only on \(\eta\shell/\tau\).  On the other
hand, they do significantly effect the question of whether a source
has a sufficiently high radiation parameter \(\Xi\) to possibly be a
dust wave.

@@

\subsubsection{Shell gas temperature}
\label{sec:shell-gas-temp}
For bows around O~stars, the shell temperature should be close to the
photoionization equilibrium value of \(\approx \SI{e4}{K}\), since the
post-shock cooling length is short in ambient densities above
\SI{0.1}{cm^{-3}} (\S~3.2 of Paper~I)
% XREF: Paper I \ref{sec:radi-cool-lengths}
and the shell does not trap the ionization front for ambient densities
below \SI{e4}{cm^{-3}} (\S~3.1 of Paper~I).
% XREF: Paper I \\ref{sec:trapp-ioniz-front}
For B~stars, on the other hand, these two density limits move closer
together (cf.~the smaller gap between the blue and red lines in
% XREF: Paper I \ref{fig:zones-v-n-plane}a
% XREF: Paper I \ref{fig:B-supergiant}
% XREF: Paper I \ref{fig:zones-v-n-plane}bc
Paper~I's Figs.~2a and 4, as compared with 2b and~c), making it more
likely that a bow will lie in a different temperature regime.  The
only source for which we have evidence that this occurs is LP~Ori, as
discussed in the next section.

@@

For two of the additional bows listed in Table~\ref{tab:observations},
we are forced to deviate from the default values for the shell
parameters.  For LP~Ori, the bow shell appears to be formed from
neutral gas \citep{ODell:2001c} and its relatively high \(\tau\) value
is more than sufficient to trap the weak ionizing photon output of a
B3 star.  We therefore move its point in
Figure~\ref{fig:All-sources-eta-tau} downward to reflect a temperature
of \SI{1000}{K} (or, equivalently, a magnetically supported shell with
\(v\alfven \approx \SI{3}{km.s^{-1}}\)).  For the case of the Orion
Trapezium star \thD{}, we find that using the default parameters
results in a placement well inside the forbidden zone of
Figure~\ref{fig:All-sources-eta-tau} (indicated by fainter symbol).
For this object there is no reason to suspect anything but the usual
photoionized temperature of \SI{e4}{K}, but its placement could be
resolved either by decreasing the shell thickness, or decreasing the
UV dust opacity, or both. Given the moderate limb brightening seen in
the highest resolution images of the Ney--Allen nebula
\citep{Robberto:2005a, Smith:2005a}, the shell thickness is unlikely
to be less than half our default value.  But, if this were combined
with a factor 5 decrease in \(\kappa\), as suggested by
\citet{Salgado:2016a}, then this would be sufficient to move the
source up to the RBW line, or slightly above.

@@

K18 derive mass loss rates for their sources using a method that is
different from the one that we employ above.  Both methods are based
on determining the stellar wind ram pressure that supports the bow
shell, but K18 do so via the following steps:
\begin{enumerate}[K1.]
\item \label{K1} The line-of-sight mass column through the shell is
  calculated by combining the peak surface brightness at \SI{70}{\um},
  \(S_{70}\), with a theoretical emissivity per nucleon,
  \(j_{70}(U)\), from \citet{Draine:2007a}:
  \(\Sigma\LOS = S_{70} / \bar{m} j_{70}(U)\).  This depends on
  knowledge of the stellar radiation field at the shell:
  \(U \propto L_* / R_0^2\).
\item \label{K2} The shell density is found from the line-of-sight
  mass column using an observationally determined ``chord diameter'',
  \(\ell\), which is assumed to be equal to the depth along the line
  of sight: \(\rho\shell = \Sigma\LOS / \ell\).
\item \label{K3} The internal ram pressure is equated to the external
  ram pressure, which is found by assuming a stream velocity of
  \SI{30}{km.s^{-1}} and a compression factor of 4 across the outer
  shock:
  \(P_{\text{stream}} = 0.25 \rho\shell \times
  (\SI{30}{km.s^{-1}})^2\).
\end{enumerate}
There are clear parallels but also differences between steps
K\ref{K1}--K\ref{K3} and our own steps P\ref{P1}--P\ref{P3}.  Our
step~P\ref{P1} depends on the total observed infrared flux of the bow
combined with an assumption about the grain opacity at ultraviolet
wavelengths, while step~K\ref{K1} depends on the peak brightness at a
single wavelength combined with an assumption about the grain
emissivity at infrared wavelengths.  Our step~P\ref{P2} requires an
assumption about the relative thickness of the shell, while
step~K\ref{K2} is more directly tied to observations.
On the other hand, step~K\ref{K3} makes a roughly equivalent
assumption about the shock compression factor,\footnote{%
  In reality, the compression factor may be larger or smaller than 4,
  depending on the efficiency of the post-shock cooling (see
  % XREF Paper I \ref{sec:radi-cool-lengths}
  \S~3.2 of Paper~I).  For instance, for \(v = \SI{30}{km.s^{-1}}\) as
  assumed by K18 and \(T = \SI{e4}{K}\), one has a Mach number of
  \(\M_0 = 2.63\) and a compression factor of 2.8 for a non-radiative
  shock (by
  % XREF Paper I \ref{eq:shock-n-jump}
  eq.~[27] of Paper~I) or a factor of
  \(\M_0^2 =6.9\) for a strongly radiative one.} %
and a further assumption about the stream velocity. These assumptions
are not necessary for our step~P\ref{P3}, but we do need to assume a
value for the shell gas temperature.

@@

For five of the K18 sources (380, 407, 409, 410, 411), our method of
\S~\ref{sec:mass-loss-determ} gives only an upper limit to the wind
mass-loss rate if one assumes a factor-three uncertainty in
\(\eta\shell\), raising the possibility that they may be
radiation-supported instead of wind-supported (see Paper~I).  It is
notable that these sources are among the smallest in the K18 sample,
all with \(R_0 < \SI{0.1}{pc}\).  The two Orion Nebula bows, LP~Ori
and \thD{} also fall into this category, and they are smaller still,
with \(R_0 < \SI{0.01}{pc}\).  Six of these 7 stars also occupy a
narrow range of spectral type\footnote{%
  However, the statistically significance is not strong.  For
  instance, out of the twenty K18 sources, 7 have spectral type O8 or
  earlier.  Therefore, taking as the null hypothesis that the 5
  sources were randomly selected from the 20, then the chance that
  they should all be O9 or later is
  \(p = (1 - \frac{7}{20})^5 = 0.116\), which fails to meet the
  conventional \(2\sigma\) significance level of \(p = 0.05\),
  thereby lending only weak support for rejection of the null.} %
between O9 and B0, corresponding to
\(T_{\text{eff}} = \text{\SIrange{28}{31}{kK}}\) (LP~Ori is a much
cooler B2 star with \(T_{\text{eff}} = \SI{20}{kK}\),
\citealp{Petit:2008a, Alecian:2013a}).

>>>>

A fundamental parameter is the optical depth, \(\tau\), of the bow
shell to UV radiation, which determines what fraction of the stellar
photon momentum is available to support the shell (see
\S~2.1 of Paper~I).  But the same photons also heat the
dust grains in the bow, which re-radiate that energy predominantly at
mid-infrared wavelengths (roughly \SIrange{10}{100}{\um}) with
luminosity \(L\IR\).  Assuming that Ly\(\alpha\) and mechanical
heating of the dust shell is negligible (see
\S~\ref{sec:unimp-other-heat}) and that the emitting shell subtends a
solid angle \(\Omega\), as seen from the star, then the optical depth
can be estimated as
\begin{equation}
  \label{eq:tau-empirical}
  \tau = -\ln \left( 1 - \frac{4\pi}{\Omega} \frac{L\IR}{L_*} \right)
  \approx \frac{2 L\IR}{L_*} \ ,
\end{equation}
where the last approximate equality holds if \(\tau \ll 1\) and the shell
emission covers one hemisphere.\footnote{%
  Note that the \(\tau\) of Paper~I is not exactly the same as the
  \(\tau\) of equation~\eqref{eq:tau-empirical}, but is larger by a
  factor of
  \(Q_P / Q_{\text{abs}} = 1 + \varpi (1 - g)/(1 - \varpi)\), where
  \(\varpi\) is the grain albedo and \(g\) the scattering asymmetry.
  For standard ISM grain mixtures,
  \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
  wavelengths.}
% This is plotted in Figure~6a of Paper~II for Cloudy's default ISM
% grain mixture (see \S~\ref{sec:grain-temp-emiss} below), where it
% can be seen that
% \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
% wavelengths (\(\lambda = \text{\SIrange{0.05}{0.2}{\um}}\)), so the
% correction is small.

A second important parameter is the thermal plus magnetic pressure in
the shocked shell, which is doubly useful since in a steady state it
is equal to \emph{both} the internal supporting pressure (wind ram
pressure plus absorbed stellar radiation) \emph{and} the external
confining pressure (ram pressure of ambient stream).  The shell
pressure is not given directly by the observations, but can be
determined by the following three steps:
\begin{enumerate}[P1.]
\item \label{P1} The shell mass column (\si{g.cm^{-2}}) can be
  estimated from the optical depth by assuming an effective UV
  opacity: \(\Sigma\shell = \tau / \kappa\)
\item \label{P2} The shell density (\si{g.cm^{-3}}) can be found from
  the mass column if the shell thickness is known:
  \(\rho\shell = \Sigma\shell / h\shell\).  In the absence of other information, a
  fixed fraction of the shell radius can be used.  In particular, we
  normalize by a typical value of one~quarter the star--apex distance:
  \(h_{1/4} = h\shell / (0.25 R_0)\).  This corresponds to a Mach
  number \(\M_0 = \surd 3\) if the stream shock is radiative, or
  \(\M_0 \gg 1\) if non-radiative (see \S~3.2 of Paper~I). Further
  discussion is given in \S~\ref{app:bow-shock-data} below. 
\item \label{P3} Finally, the pressure (\si{dyne.cm^{-2}}) follows by
  assuming values for the sound speed and Alfvén speed:
  \(P\shell = \rho\shell (\sound^2 + \frac12 v\alfven^2) \).
\end{enumerate}
It is natural to normalize this pressure to the stellar radiation
pressure at the shell, so we define a shell momentum efficiency
\newcommand\pc{\ensuremath{_{\text{pc}}}}
\begin{equation}
  \label{eq:eta-shell}
  \eta\shell \equiv \frac{P\shell}{P\rad}
  = \frac{4\pi R_0^2\, (\sound^2 + \frac12 v\alfven^2)\, c\, \tau}{L_*\, \kappa\, h\shell}
  \approx 245 \frac{R\pc \, T_4 \, \tau}{L_4 \, \kappa_{600} \, h_{1/4}} \ , 
\end{equation}
where \(c\) is the speed of light. In the last step we have assumed
ionized gas at temperature \(\num{e4}\,T_4\,\si{K}\) with negligible
magnetic support (\(v\alfven \ll \sound\)) and written the stellar
luminosity and shell parameters in terms of typical values, which we
summarize below.  Note that the shell momentum efficiency is simply
the reciprocal of the radiation parameter from Paper~II's
equation~(23): \(\eta\shell = \Xi\shell^{-1}\), which provides yet a third
use for \(\eta\shell\), since \(\Xi\) is paramount in determining whether
the grains and gas remain well-coupled (see \S~4.4 of Paper~II).

@@

\begin{table}
  \centering
  \caption[Observational]{Key observational parameters for star/bow systems}
  \label{tab:observations}
  \begin{tabular}{l S S S}
    \toprule
    Star & {\(L_* / \SI{e4}{L_\odot}\)}
    & {\(L_{\text{IR}} / \si{L_\odot}\)} & {\(R_0 / \si{pc}\)} \\
    \midrule
    \thD & 2.95 & 620 & 0.003 \\
    LP~Ori & 0.16 & 240 & 0.01 \\
    \(\sigma\)~Ori & 6.0 & 15 & 0.12 \\[\smallskipamount]
    K18 Sources & \numrange{1.4}{87} & \numrange{8}{2800} & \numrange{0.02}{1.35} \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The UV optical depth of the bow shell is obtained
(eq.~[\ref{eq:tau-empirical}]) from the ratio of infrared shell
luminosity to stellar luminosity.  The inverse of this ratio is given
in Table~5 of K17, but we choose to re-derive the values since the
spectral classification of some of the sources was revised between K17
and K18.  Although K17 found the total shell fluxes from fitting dust
emission models to the observed SEDs, we adopt the simpler approach of
taking a weighted sum of the flux densities \(F_\nu\) (in \si{Jy}) in
three mid-infrared bands:
\begin{multline}
  \label{eq:total-ir-flux}
  F_{\text{IR}}  \approx \bigl[  2.4\,(F_8 \text{ or } F_{12})
    + 1.6\,(F_{22} \text{ or } F_{24})  \\
  + 0.51\,F_{70}\bigr]
  \,\times \SI{e-10}{erg.s^{-1}.cm^{-2}} \ ,
\end{multline}
where \(F_8\) is Spitzer IRAC \SI{8.0}{\um}, \(F_{24}\) is Spitzer
MIPS \SI{23.7}{\um}, \(F_{12}\) and \(F_{22}\) are WISE bands 3 and 4,
and \(F_{70}\) is Herschel PACS \SI{70}{\um}.  The weights are chosen
so that the integral \(\int_0^\infty F_\nu \,d\nu\) is approximated by the
quadrature sum \(\Sigma_k F_k\, \Delta\nu_k\), under the assumption that fluxes in
shorter (e.g., IRAC \SI{5.8}{\um}) and longer (e.g., PACS
\SI{150}{\um}) wavebands are negligible.  Shell fluxes are converted
to luminosities using the assumed distance to each source, and stellar
luminosities are taken directly from K18 Table~2, based on
spectroscopic classification and the calibrations of gravity and
effective temperature from \citet{Martins:2005a}.

@@

The fundamental observational quantities that go into determining
\(\tau\) and \(\eta\shell\) for each source are distance, \(D\); stellar
luminosity, \(L_*\); total infrared flux, \(F_{\text{IR}}\); and bow
angular apex distance, \(\theta\).  From these, the shell radius and
infrared luminosity are found as \(R_0 = \theta D\) and
\(L_{\text{IR}} = 4\pi D^2 F_{\text{IR}}\).  Rather than clutter the
diagram with error bars, we instead show the sensitivity to
observational errors in the lower-right box, where each arrow
corresponds to a factor of two increase (0.3~dex) in each quantity:
\(D\), \(L_*\), \(F_{\text{IR}}\), and \(\theta\).  We now calculate
uncertainty estimates for individual observational quantities that are
used in deriving not only \(\tau\) and \(\eta\shell\) but also mass-loss
rates, as discussed later in \S~\ref{sec:stellar-wind-mass}.

@@

Assuming that the uncertainty in each observational quantity is
independent, we can now combine them using the techniques described in
Appendix~\ref{sec:comb-uncert-covar} to find the \(\pm 1~\sigma\) error
ellipse, shown in blue in the figure.  It can be seen that
observational uncertainties in \(\tau\) and \(\eta\shell\) are highly
correlated: the dispersion is \SI{0.7}{dex} in the product
\(\eta\shell \tau\) but only \SI{0.16}{dex} in the ratio
\(\eta\shell/\tau\), with stellar luminosity errors dominating in both
cases.  Observational uncertainties are therefore relatively
unimportant in determining whether a given source is wind-driven or
radiation-driven, which depends only on \(\eta\shell/\tau\).  On the other
hand, they do significantly affect the question of whether a source
has a sufficiently high radiation parameter \(\Xi\) to possibly be a
dust wave.

@@

\subsubsection{Shell gas temperature}
\label{sec:shell-gas-temp}
For bows around O~stars, the shell temperature should be close to the
photoionization equilibrium value of \(\approx \SI{e4}{K}\), since the
post-shock cooling length is short in ambient densities above
\SI{0.1}{cm^{-3}}
% (\S~3.2 of Paper~I)
% XREF: Paper I \ref{sec:radi-cool-lengths}
and the shell does not trap the ionization front for ambient densities
below \SI{e4}{cm^{-3}} (see Paper~I's \S\S~3.1 and 3.2 for details).
% (\S~3.1 of Paper~I).
% XREF: Paper I \\ref{sec:trapp-ioniz-front}
For B~stars, on the other hand, these two density limits move closer
together,
% (cf.~the smaller gap between the blue and red lines in
% XREF: Paper I \ref{fig:zones-v-n-plane}a
% XREF: Paper I \ref{fig:B-supergiant}
% XREF: Paper I \ref{fig:zones-v-n-plane}bc
% Paper~I's Figs.~2a and 4, as compared with 2b and~c),
making it more likely that a bow will lie in a different temperature
regime.  The only source for which we have evidence that this occurs
is LP~Ori, as discussed in the next section.

@@

For two of the additional bows listed in Table~\ref{tab:observations},
we are forced to deviate from the default values for the shell
parameters.  For LP~Ori, the bow shell appears to be formed from
neutral gas \citep{ODell:2001c} and its relatively high \(\tau\) value is
more than sufficient to trap the weak ionizing photon output of a B3
star.  We therefore move its point in
Figure~\ref{fig:All-sources-eta-tau} downward by a factor of ten,
which could be a thermally supported neutral shell at \SI{2000}{K} or
a magnetically supported shell with
\(v\alfven \approx \SI{3}{km.s^{-1}}\)).  For the case of the Orion
Trapezium star \thD{}, we find that using the default parameters
results in a placement well inside the forbidden zone of
Figure~\ref{fig:All-sources-eta-tau} (indicated by fainter symbol).
For this object there is no reason to suspect anything but the usual
photoionized temperature of \SI{e4}{K}, but its placement could be
resolved either by decreasing the shell thickness, or decreasing the
UV dust opacity, or both. Given the moderate limb brightening seen in
the highest resolution images of the Ney--Allen nebula
\citep{Robberto:2005a, Smith:2005a}, the shell thickness is unlikely
to be less than half our default value.  But, if this were combined
with a factor 5 decrease in \(\kappa\), as suggested by
\citet{Salgado:2016a}, then this would be sufficient to move the
source up to the RBW line, or slightly above.

@@

K18 derive mass-loss rates for their sources using a method that is
different from the one that we employ above.  Both methods are based
on determining the stellar wind ram pressure that supports the bow
shell, but K18 do so via the following steps:
\begin{enumerate}[K1.]
\item \label{K1} The line-of-sight mass column through the shell is
  calculated by combining the peak surface brightness at \SI{70}{\um},
  \(S_{70}\), with a theoretical emissivity per nucleon,
  \(j_{70}(U)\), from \citet{Draine:2007a}:
  \(\Sigma\LOS = S_{70} / \bar{m} j_{70}(U)\).  This depends on
  knowledge of the stellar radiation field at the shell:
  \(U \propto L_* / R_0^2\).
\item \label{K2} The shell density is found from the line-of-sight
  mass column using an observationally determined ``chord diameter'',
  \(\ell\), which is assumed to be equal to the depth along the line
  of sight: \(\rho\shell = \Sigma\LOS / \ell\).
\item \label{K3} The internal ram pressure is equated to the external
  ram pressure, which is found by assuming a stream velocity of
  \SI{30}{km.s^{-1}} and a compression factor of 4 across the outer
  shock:
  \(P_{\text{stream}} = 0.25 \rho\shell \times
  (\SI{30}{km.s^{-1}})^2\).
\end{enumerate}
There are clear parallels but also differences between steps
K\ref{K1}--K\ref{K3} and our own steps P\ref{P1}--P\ref{P3}.  Our
step~P\ref{P1} depends on the total observed infrared flux of the bow
combined with an assumption about the grain opacity at ultraviolet
wavelengths, while step~K\ref{K1} depends on the peak brightness at a
single wavelength combined with an assumption about the grain
emissivity at infrared wavelengths.  Our step~P\ref{P2} requires an
assumption about the relative thickness of the shell, while
step~K\ref{K2} is more directly tied to observations.
On the other hand, step~K\ref{K3} makes a roughly equivalent
assumption about the shock compression factor,\footnote{%
  In reality, the compression factor may be larger or smaller than 4,
  depending on the efficiency of the post-shock cooling (see
  % XREF Paper I \ref{sec:radi-cool-lengths}
  \S~3.2 of Paper~I).  For instance, for \(v = \SI{30}{km.s^{-1}}\) as
  assumed by K18 and \(T = \SI{e4}{K}\), one has a Mach number of
  \(\M_0 = 2.63\) and a compression factor of 2.8 for a non-radiative
  shock (by
  % XREF Paper I \ref{eq:shock-n-jump}
  eq.~[27] of Paper~I) or a factor of
  \(\M_0^2 =6.9\) for a strongly radiative one.} %
and a further assumption about the stream velocity. These assumptions
are not necessary for our step~P\ref{P3}, but we do need to assume a
value for the shell gas temperature.

@@

For five of the K18 sources (380, 407, 409, 410, 411), our method of
\S~\ref{sec:mass-loss-determ} gives only an upper limit to the wind
mass-loss rate if one assumes a factor-three uncertainty in
\(\eta\shell\), raising the possibility that they may be
radiation-supported instead of wind-supported (see Paper~I).  It is
notable that these sources are among the smallest in the K18 sample,
all with \(R_0 < \SI{0.1}{pc}\).  The two Orion Nebula bows, LP~Ori
and \thD{} also fall into this category, and they are smaller still,
with \(R_0 < \SI{0.01}{pc}\).  Six of these 7 stars also occupy a
narrow range of spectral type\footnote{%
  However, the statistical significance is not strong.  For
  instance, out of the twenty K18 sources, 7 have spectral type O8 or
  earlier.  Therefore, taking as the null hypothesis that the 5
  sources were randomly selected from the 20, then the chance that
  they should all be O9 or later is
  \(p = (1 - \frac{7}{20})^5 = 0.116\), which fails to meet the
  conventional \(2\sigma\) significance level of \(p = 0.05\),
  thereby lending only weak support for rejection of the null.} %
between O9 and B0, corresponding to
\(T_{\text{eff}} = \text{\SIrange{28}{31}{kK}}\) (LP~Ori is a much
cooler B2 star with \(T_{\text{eff}} = \SI{20}{kK}\),
\citealp{Petit:2008a, Alecian:2013a}).
